---
title: "programmer - proj.4"
author: "Simran Makwana"
date: "4/15/2020"
output: word_document
---

```{r echo = FALSE, message = FALSE}
#install.packages("Seurat")
library(Seurat)
#BiocManager::install("tximport")
library(tximport) # must be version >= 1.12.2

# path to *our output directory*
files <- file.path("../umi_count_files/GSM2230758/GSM2230758_alevin_4/alevin/quants_mat.gz")

# check
file.exists(files)

# Reading in the alevin quants quants
txi <- tximport(files, type="alevin")

```

Filter out low-quality cells and genes. You should consider the minimum number of non-zero count genes per cell, the number of non-zero count cells per gene, etc. NB: The UMI counts matrix has Ensembl gene identifiers, but it may be helpful to map them to gene symbols. There are many ways to do this.
```{r echo = FALSE, message = FALSE}
# map ensembl gene identifiers to gene symbols
#BiocManager::install("SeqGSEA")
library(SeqGSEA)

# no filtering
panc <- CreateSeuratObject(counts = txi$counts, project = "panc") 
#panc <- CreateSeuratObject(counts = txi_new$counts, project = "panc") 

print("no filtering")
panc

# filter out genes that are detected in less than 3 cells, and cells that are detected in less than 200 features
panc <- CreateSeuratObject(counts = txi$counts, project = "panc", min.cells = 3, min.features = 20)
print("filters: min.cells = 3, min.features = 20")
panc

# low mito
genes <- substr(panc@assays$RNA@counts@Dimnames[[1]], start = 1, stop = 15)
conversion <- convertEnsembl2Symbol(genes)
genes <- conversion$hgnc_symbol

panc@assays$RNA@counts@Dimnames[[1]] <- genes
panc@assays$RNA@data@Dimnames[[1]] <- genes

panc[["percent.mt"]] <- PercentageFeatureSet(panc, pattern = "^MT-")
VlnPlot(panc, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
plot1 <- FeatureScatter(panc, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(panc, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))

#duplicts...
# filter out cells with too few (low quality or empty droplets) or too many (doublets or multiplets) genes
panc <- subset(panc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)

# report
print("filters: nFeature_RNA>200 and nFeature_RNA < 2500")
panc

```

Filter out low variance genes. Further filter the counts matrix to include only highly variable features that are likely to be informative. Choose an appropriate number of features based on the data. Recall that the counts matrix must be normalized before comparing cells to one another.
```{r echo = FALSE, message = FALSE}
# normalize counts matrix
panc <- NormalizeData(panc)

panc[["RNA"]]@meta.features <- data.frame(row.names = rownames(panc[["RNA"]]))

# filter out low variance genes; only include highly variable features (top 2000)
panc <- FindVariableFeatures(panc, selection.method = "vst", nfeatures = 2000)

# scale genes 
panc <- ScaleData(panc)

# report
print("normalized, filtered out low variable genes, scaled")
panc

# change var.features to gene names for analyst
#genes <- panc@assays$RNA@counts@Dimnames[[1]]
#varnames <- panc@assays$RNA@var.features

# save object
save(panc, file = "../output/processed_panc.rda")

# genes the analyst needs
needed_genes <- c("GCG", "INS", "SST", "PPY", "GHRL", "KRT19", "CPA1", "PDGFRB", "VWF", "PECAM1", "CD34", "CD163", "CD68", "IgG", "CD3", "CD8", "TPSAB1", "KIT", "CPA3")


#mygenes <- substr(panc@assays$RNA@counts@Dimnames[[1]], start = 1, stop = 15)
#conversion <- convertEnsembl2Symbol(mygenes)
#mygenes <- conversion$hgnc_symbol

#panc@assays$RNA@counts@Dimnames[[1]] <- mygenes

print("the following genes are not in the filtered list:")
needed_genes[which(!needed_genes %in% panc@assays$RNA@counts@Dimnames[[1]])]
```

Identify clusters of cell type subpopulations. Use a clustering method of your choosing to discover clusters of cells. Identify the number of cells in each cluster.
```{r echo = FALSE, message = FALSE}
# dimension reduction (PCA)

panc <- RunPCA(panc, features = VariableFeatures(object = panc))

# find clusters
ElbowPlot(panc)

panc <- FindNeighbors(panc, dims = 1:5i)
panc <- FindClusters(panc, resolution = 1.7)
cluster_assignments <- Idents(panc)
plot(cluster_assignments)

save(panc, file = "../output/processed_panc_with_clusters.rda")


```



